Impact ionization processes in a photodriven Mott insulator: influence of phononic dissipation

### 2022-08-23 Paper1.tex
Added sources for figures
Final version 23.08.2022

\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{bm}
\usepackage{cancel}
\usepackage{xspace}
\usepackage{braket}
\usepackage{xcolor}   
\usepackage{calrsfs}
\usepackage{comment}
\usepackage{placeins}%enables \FloatBarier
\usepackage{hyperref}

@@

%% REFERENCE
\newcommand{\fig}[1]{Fig.\thinspace{}\ref{#1}}
\newcommand{\figs}[1]{Figs.\thinspace{}\ref{#1}} 
\newcommand{\Fig}[1]{Figure \ref{#1}}
\newcommand{\eq}[1]{Eq.\thinspace{}(\ref{#1})}
\newcommand{\Eq}[1]{Eq.\thinspace{}(\ref{#1})}
\newcommand{\eqs}[1]{Eqs.\thinspace{}(\ref{#1})}
\newcommand{\Eqs}[1]{Eqs.\thinspace{}(\ref{#1})}
\newcommand{\feqs}{eqs.\@\xspace}
\newcommand{\fEqs}{Eqs.\@\xspace}
\newcommand{\ch}{Ch.\@\xspace}
\newcommand{\Ch}{Ch.\@\xspace}
\newcommand{\pa}{par.\@\xspace}
\newcommand{\Pa}{Par.\@\xspace}
\newcommand{\se}{Sec.\@\xspace}
\newcommand{\Se}{Sec.\@\xspace}
\newcommand{\app}{App.\@\xspace}
\newcommand{\App}{App.\@\xspace}

@@

%Defs
\newcommand{\und}{\underline}
\newcommand{\iim}{\Im{}\,}
\newcommand{\li}{\mathcal L}
\newcommand{\ga}[1]{\Gamma^{(#1)}}
\newcommand{\up}{\ensuremath{\uparrow}}
\newcommand{\dw}{\ensuremath{\downarrow}}
\newcommand{\vv}[1]{\boldsymbol{#1}}
\newcommand{\Dph}{\underline{\Delta}_\mathrm{ph}(\omega)}
\newcommand{\Daux}{\underline{\Delta}_\mathrm{aux}(\omega)}
\newcommand{\DRph}{\Delta^R_\mathrm{ph}(\omega)}
\newcommand{\DRaux}{\Delta^R_\mathrm{aux}(\omega)}
\newcommand{\DKph}{\Delta^K_\mathrm{ph}(\omega)}
\newcommand{\DKaux}{\Delta^K_\mathrm{aux}(\omega)}

%choose
%\newcommand{\phys}{\text{phys}}
%\newcommand{\phys}{\text{DMFT}}
\newcommand{\phys}{} %my preference

@@

\definecolor{orange}{RGB}{252,77,6}
\definecolor{brown}{RGB}{200,127,50}

@@

@@

\title{Impact ionization processes in driven Mott insulating layer: influence of fermion baths and phonons}

@@

We study a model for photovoltaic energy collection consisting of a Mott insulating layer coupled to local acoustic phonons between two wide-band fermion baths at different chemical potentials, driven to the non-equilibrium steady state by an external AC field. We treat electron correlations with non-equilibrium dynamical mean field theory (DMFT) using the so-called auxiliary master equation approach as impurity solver, incuding phonons via the Migdal approximation.

We show that with a small coupling to fermion baths, the peak of the photocurrent is mainly due to impact ionization processes, while with higher coupling is due primarily to direct excitations. This occurs also for higher electric field amplitudes, that boost overall the photocurrent and the double occupation. Acoustic phonons slightly enhance the photocurrent for small driving frequencies and suppress it at higher values, in correspondence to the peak, disregarding the dominant electronic scattering process taking place. Their effect is independent of the bath-layer coupling value and the electric field amplitude.

@@

\pgcomm{IS FERMION BATHS THE APPROPRIATE TERM?}

The idea of designing photovoltaic devices exploiting the Mott gap to convert electromagnetic radiation into energy has become popular in the recent years \cite{mano.10,li.ch.13,gu.gu.13,co.ma.14,wa.li.15}. In particular, it has been
suggested that in strongly correlated materials, highly excited charge carriers could use their extra energy to excite additional carriers across the Mott gap via impact ionization \cite{mano.10,co.ma.14}, thus
potentially improving their efficiency beyond the Shockley-Queisser limit \cite{sh.qu.61}. Although impact ionization is also present in conventional semiconductor devices, the time scales for electron-electron scattering are typically much longer than in correlated materials, so that highly excited electrons will mostly dissipate their energy to phonons.

Experimentally, evidence for fast carrier multiplication processes has indeed been detected via pump-probe experiments in VO$_2$ \cite{ho.bi.16}. Oxide heterostructures based on LaVO$_3$/SrTiO$_3$ have been identified as promising candidates, due to the ideal band gap and the strong polar field, able to separate the excited charge carriers \cite{as.bl.13}. There are however some drawbacks, such as the low mobility of the carriers \cite{wa.li.15,je.re.18}, which still question these materials’ applicability as efficient solar cells. In this regard, it has to be mentioned that, while a large-scale application of Mott-based solar cells may be hard to achieve, the employment of such systems as photodetectors may be more promising on the long run, due to their high photoresponsivity.

Therefore, the properties of the Mott-based photovoltaic mechanism deserve to be investigated further, from the scientific point of view and for future applications in order to explore, whether alternative unexpected paths could be opened. And in fact, a lot of theoretical work has been done, in order to understand the impact ionization process (see, e.g. \cite{co.ma.14,ec.we.11,ec.we.13,we.he.14,pe.be.19,so.do.18,ka.wo.20,ma.ev.22}).

The effects of the coupling to metallic leads and of phonons dissipation are essential features to properly establish the occurrence of impact ionization in Mott-based photovoltaic setups. They haven't been taken into account up to now and in this article we address such issues.

% First, because the effect of phonons on impact ionization is still to clarify completely and therefore there are still questions regarding the physical mechanisms to be answered. Once such questions are answered, a more detailed treatment of phonons, together with DFT may address more accurately the precise entity of the their effect.

\begin{figure}[ht]
% \captionsetup{width=\textwidth} 
\includegraphics[width=0.9\linewidth]{./figures_Paper1/setup.pdf}
\caption{(Colour online) Schematic representation of the considered setup. A correlated layer with local Hubbard interaction $U$ and electron-phonon coupling $g$ (green) is sandwiched between two non-interacting layers (red) via the coupling $v$ (black). The latter are in turn coupled via $\Gamma$ (orange) to wide-band fermion reservoirs (blue) at different chemical potentials $\mu_{l/r}$.}
\label{fig:setup}
\end{figure}

The setup used is depicted in Fig.\ref{fig:setup}: a Hubbard insulating layer with local acoustic phonons under AC driving is put between two intermediate non-interacting layers coupled to wide-band fermion reservoirs at different chemical potentials, acting as metallic leads. For sake of simplicity, we refer for the rest of the article to the non-interacting layers and fermion reservoirs together simply as fermion baths. We address the non-equilibrium steady state situation under an external AC field, treated via nonequilibrium Floquet dynamical mean-field theory (F-DMFT).

Our goal is to show for which values of the bath-layer coupling impact ionization clearly occurs and how is influenced by the effect of phonons.

We limit our analysis to qualitative and semi-quantitative features of the influence of fermion baths and phonons on photocurrent, double occupation and spectral properties. 
An accurate quantitative analysis of such effects and their relative signatures in possible experimental setups of Mott solar cells is outside the scope of this work.

We find that a small coupling to fermion baths clearly enhances impact ionization respect to a bigger coupling, for which direct excitations are dominant. Impact ionization therefore plays a crucial role, compensating the smaller charge injection and giving the same magnitude in the photocurrent of the case with a bigger coupling. Such results remain the same considering higher values of electric field amplitude, although the current and double occupation are greatly increased. The phonons slightly enhance the photocurrent for small driving frequencies, while suppressing it at higher values, in corrispondence to the peak, disregarding the dominant electron scattering process taking place. Such influence of phonons on the photocurrent is substantially unaltered sweeping over different values of the bath-layer coupling and of electric field amplitude.

The structure of the paper is the following: in Sec.\ref{sec:Model} we describe the model Hamiltonian. Methods, formalism and observables used are explained in Sec.\ref{sec:Method_formalism}. We discuss our results in Sec.\ref{sec:results}. We present our conclusions in Sec.\ref{sec:conclusions}.

@@

Our setup (Fig.\ref{fig:setup}) consists of a central Hubbard layer at half filling with local phonon coupling, under an external AC driving,attached to fermionic baths:

\begin{equation}
\label{eq:Hamiltonian}
\begin{split} 
\hat{H}(t)&=\varepsilon_c \sum_{i\sigma}\hat{n}_{i\sigma} -\sum_{\sigma}\sum_{(i,j)} t_{ij}(t) \hat{c}^{\dagger}_{i\sigma} \hat{c}_{j\sigma}
\\ & + U \sum_{i} \hat{n}_{i\uparrow} \hat{n}_{i\downarrow}+\hat{H}_{e-ph} + \hat{H}_{ph}+\hat{H}_{bath}
\end{split}
\end{equation}

The operator $\hat{c}^{\dagger}_{i\sigma}$ ($\hat{c}_{i\sigma}$) creates (annihilates) an electron with spin $\sigma= \lbrace \uparrow,\downarrow \rbrace$ on the \emph{i-th} lattice site, with $\hat{n}_{i\sigma}\equiv \hat{c}^{\dagger}_{i\sigma} \hat{c}_{i\sigma}$. We denote with $(i,j)$ the sum over next neighbour sites and with $\varepsilon_c \equiv -U/2$ the electron onsite energy. The AC field consists of a time-periodic, homogeneous and monochromatic electric field with frequency $\Omega$. It enters via the Peierl's substitution in the time-dependent hopping in Eq. (\ref{eq:Hamiltonian}) \cite{peie.33}:
\begin{equation}\label{eq:peierls} 
t_{ij}(t) = t_{c} \ e^{-i \frac{q}{\hbar} \left( \vec{r}_j - \vec{r}_i \right) \cdot \vec{A}(t)}, 
\end{equation} 
where $t_{c}$ is the intra-lattice hopping, $\vec{A}$(t) is the time-dependent vector potential, $\hbar$ the Planck's constant and $q$ the charge of the electron. Here $\vec{A}(t)=\vec{e}_{0} A(t)$ is along the lattice body diagonal, i.e. $\vec{e}_{0}=(1,1,\dots,1)$. For the AC field, $A(t)=\frac{\hbar}{qa}A\sin\Omega t$, with $A=-\frac{qE_0a}{\hbar\Omega}$, being $E_0$ the electric field amplitude and $a$ the lattice spacing ~\cite{ts.ok.08}. Therefore in the temporal gauge $\vec{E}= -\partial_{t}\vec{A}(t) = \vec{e}_{0}E_0 \cos\Omega t$.

The local electron-phonon coupling is included as in Ref.\cite{ma.ga.22u}, considering an acoustic phonon branch attached to each lattice site via the Hamiltonian 
\begin{equation}\label{eq:e-ph_ham} 
\hat{H}_{e-ph} = g \sum_{i\sigma} \hat{n}_{i\sigma} \hat{x}_{i},  
\end{equation}
with $g$ the electron-phonon coupling and $\hat{x}_{i} \equiv \frac{1}{\sqrt{2}} \left( \hat{b}^{\dagger}_{i} + \hat{b}_i \right)$, where $\hat{b}^{\dagger}_{i}$ ($\hat{b}_{i}$) creates (annihilates) a phonon of the acoustic branch at site $i$, with dispersion relation in $\hat{H}_{ph}$ \cite{ma.ga.22u}.

Details on the baths are provided in \ref{sec:Dyson_equation}.

In this paper we consider the correlated system as a $d$-dimensional layer in $d+1$ dimensions and we take the limit of infinite dimensions $d \rightarrow \infty$ by rescaling the hopping as $t_{c}=t^{*}/2\sqrt{d}$ \footnote{We tune $t^*$ such that we reproduce as best as possible the DOS of a 2D layer with $t_c$ = 1, following Ref.\cite{so.do.18}.}. With $a=1$, sums over crystal momentum ${\bf k}$ of a generic quantity $\chi$ read $\sum_{{\bf k}} \chi(\omega,{\bf k}) \rightarrow \int d\epsilon \int   d\overline{\epsilon} \ \rho(\epsilon,\overline{\epsilon}) \chi(\omega;\epsilon,\overline{\epsilon})$ with $\rho(\epsilon,\overline{\epsilon}) = (1/\pi t^{* 2}) \ exp(-( \epsilon^{2} + \overline{\epsilon}^{2})/t^{* 2})$ the joint density of states (JDOS) ~\cite{ts.ok.08}, with  
\begin{align}\label{eq:d-dim_crystal_dep}
\begin{split}  
\epsilon & = -2t_{c} \sum_{i=1}^{d} cos(k_i a) \\  
\overline{\epsilon}& = -2t_{c}\sum_{i=1}^{d} sin(k_i a). \\
\end{split}  
\end{align}

In this paper we set $\hbar = k_B = a = 1 = -q$, so that every physical quantity is in units of $t^{*}$.

@@

To correctly describe the nonequilibrium periodic steady-state (NESS), we use the Floquet generalization of non-equilibrium Green's function (NEGF) approach \cite{ts.ok.08}. Every function $G(t;t{'})$ satisfying the periodicity relation $G(t;t{'})=G(t+\tau;t{'}+\tau)$, with $\tau=2\pi/\Omega$ the period related to the external driving frequency $\Omega$, may be represented as \cite{ts.ok.08,sc.mo.02u,jo.fr.08}

\begin{equation}
\label{eq:FloquetGF} 
\underline{G}_{mn}(\omega) =\int dt_{rel} \int_{-\tau/2}^{\tau/2} \frac{dt_{av}}{\tau} e^{i[\left(\omega+m\Omega\right) t -\left( \omega+n\Omega\right)t{'}]} \underline{G}(t,t'),
\end{equation}

known as the the \emph{Keldysh-Floquet} GF, and

\begin{equation}\label{eq:WignerGF}
\underline{G}_{l}(\omega) =\int dt_{rel} \int_{-\tau/2}^{\tau/2} \frac{dt_{av}}{\tau} e^{i\omega t_{rel} + il\Omega t_{av}} \underline{G}(t,t'),
\end{equation}

the \emph{Wigner} GF ~\cite{ts.ok.08}. $t_{rel} = t-t{'}$ and $t_{av} = (t+t{'})/2$ are the relative and average-time variables. Eqs. (\ref{eq:FloquetGF}) and (\ref{eq:WignerGF}) are related via the relation

\begin{equation}\label{eq:Wig2Fl}
\underline{G}_{mn}(\omega)=\underline{G}_{m-n}\left(\omega+\frac{m+n}{2}\Omega \right),
\end{equation}

We denote in the rest of this work a Floquet-represented matrix as either $X_{mn}$, with the explicit indices, or using a boldface letter, $\bf{X}$. A single index as subscript, $X_{l}$, is reserved for the Wigner representation. The underline indicates the overall structure 
\begin{equation}\label{eq:Keld-structure}
\underline{{\bf G}} \equiv 
\begin{pmatrix}
{\bf G}^{R} & {\bf G}^{K}\\
{\bf 0}         & {\bf G}^{A} \\
\end{pmatrix},
\end{equation}

which contains the \emph{retarded}, \emph{advanced} and \emph{Keldysh} components $\textbf{G}^{\text{R,A,K}}$, where $\textbf{G}^{\text{A}}=(\textbf{G}^{\text{R}})^{\dagger}$. The \emph{Keldysh} component is defined as $\textbf{G}^{K} \equiv \textbf{G}^{>} + \textbf{G}^{<}$, with $\textbf{G}^{\lessgtr}$ the \emph{lesser} and \emph{greater} components~\cite{schw.61,keld.65,ra.sm.86,ha.ja}

@@

\begin{equation}\label{eq:FullDysonEq}
\underline{{\bf G}}^{-1}(\omega_n;\epsilon,\overline{\epsilon}) = \underline{{\bf G}}^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon}) - \underline{{\bf \Sigma}}(\omega_n;\epsilon,\overline{\epsilon}) - \underline{{\bf \Sigma}}_{e-ph}(\omega_n;\epsilon,\overline{\epsilon}),
\end{equation}

The lattice GF of the non-interacting part of the Hamiltonian, Eq. (\ref{eq:Hamiltonian}), is

\begin{equation}\label{eq:non-int_InvGF} 
\underline{G}^{-1}_{0,mn}(\omega_n;\epsilon,\overline{\epsilon}) =  \underline{g}^{-1}_{0,mn}(\omega_n;\epsilon,\overline{\epsilon}) - \sum_{\rho\in\{l,r\}}v^2_{\rho}\underline{g}_{b,\rho}(\omega_n;\epsilon)\delta_{mn},
\end{equation}

 with $v_{l/r}$ the bath-layer coupling, $\omega_n \equiv \omega + n\Omega$, $n\in\mathbbm{Z}$, and

 \begin{align}\label{eq:inv_non-int_lat_GF_comps}
\begin{split}
\left[ g^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon})) \right]^{R}_{mn} & = \left( \omega_n + i0^{+} -\varepsilon_c \right)\delta_{mn} - \varepsilon_{mn}(\epsilon,\overline{\epsilon})) \\
\left[ g^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon})) \right]^{K}_{mn} & = 0 \\
\end{split}.
\end{align}

In the steady-state, the inverse Keldysh component is negligible \cite{so.do.18}. The Floquet dispersion relation, $\varepsilon_{mn}$, for the AC field in a hyper-cubic lattice is \cite{ts.ok.08}

>>>>

\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{bm}
\usepackage{cancel}
\usepackage{xspace}
\usepackage{braket}
\usepackage{xcolor}   
\usepackage{calrsfs}
\usepackage{comment}
\usepackage{placeins}%enables \FloatBarrier
\usepackage[colorlinks=true]{hyperref} %hidelinks=true

@@

%% REFERENCE
\newcommand{\fig}[1]{Fig.~\ref{#1}}
\newcommand{\figs}[1]{Figs.~\ref{#1}} 
\newcommand{\Fig}[1]{Figure~\ref{#1}}
\newcommand{\eq}[1]{Eq.~\eqref{#1}}
\newcommand{\Eq}[1]{Eq.~\eqref{#1}}
\newcommand{\eqs}[1]{Eqs.~\eqref{#1}}
\newcommand{\Eqs}[1]{Eqs.~\eqref{#1}}
\newcommand{\feqs}{eqs.\@\xspace}
\newcommand{\fEqs}{Eqs.\@\xspace}
\newcommand{\ch}{Ch.\@\xspace}
\newcommand{\Ch}{Ch.\@\xspace}
\newcommand{\pa}{par.\@\xspace}
\newcommand{\Pa}{Par.\@\xspace}
\newcommand{\se}{Sec.\@\xspace}
\newcommand{\Se}{Sec.\@\xspace}
\newcommand{\app}{App.\@\xspace}
\newcommand{\App}{App.\@\xspace}

@@

%Defs
\newcommand{\und}{\underline}
\DeclareMathOperator*{\rre}{Re}
\DeclareMathOperator*{\iim}{Im}
\DeclareMathOperator*{\Tr}{Tr}
\newcommand{\li}{\mathcal L}
\newcommand{\ga}[1]{\Gamma^{(#1)}}
\newcommand{\up}{\ensuremath{\uparrow}}
\newcommand{\dw}{\ensuremath{\downarrow}}
\newcommand{\vv}[1]{\boldsymbol{#1}}
\newcommand{\Dph}{\underline{\Delta}\text{ph}(\omega)}
\newcommand{\Daux}{\underline{\Delta}_\text{aux}(\omega)}
\newcommand{\DRph}{\Delta^\text{R}_\text{ph}(\omega)}
\newcommand{\DRaux}{\Delta^\text{R}_\text{aux}(\omega)}
\newcommand{\DKph}{\Delta^\text{K}_\text{ph}(\omega)}
\newcommand{\DKaux}{\Delta^\text{K}_\text{aux}(\omega)}

%choose
%\newcommand{\phys}{\text{phys}}
%\newcommand{\phys}{\text{DMFT}}
\newcommand{\phys}{} %my preference 
\newcommand{\aamp}{\mathcal{A}}

@@

\definecolor{orange}{RGB}{252,77,6}
\definecolor{brown}{RGB}{200,127,50}
\definecolor{blue}{RGB}{0,0,255}

@@

\newcommand{\jlcomm}[1]{{\color{orange} [JL COMMENT: #1]}}
\newcommand{\jlch}[1]{{\color{orange} #1}}

\newcommand{\tmcomm}[1]{{\color{blue} [TM COMMENT: #1]}}
\newcommand{\tmch}[1]{{\color{blue} #1}}

@@

%EA
\title{Impact ionization processes in a photodriven Mott insulator: influence of phononic dissipation}

@@

We study a model for photovoltaic energy collection consisting of a Mott insulating layer in presence of acoustic phonons, coupled to two wide-band fermion leads at different chemical potentials and driven by a periodic electric field. We treat electron correlations with nonequilibrium dynamical mean-field theory (DMFT) using the so-called auxiliary master equation approach as impurity solver and include dissipation by acoustic phonons via the Migdal approximation. For a small hybridization to the leads, we obtain a peak in the photocurrent as a function of the driving frequency which can be associated with impact ionization processes.
For larger hybridizations the shallow peak suggests a suppression of impact ionization with respect to direct photovoltaic excitations. Acoustic phonons slightly enhance the photocurrent for small driving frequencies and suppress it at frequencies around the main peak. This effect occurs at all considered hybridization strengths.

@@

The idea of designing photovoltaic devices exploiting the Mott gap to convert electromagnetic radiation into energy has become popular in recent years~\cite{mano.10,li.ch.13,gu.gu.13,co.ma.14,wa.li.15}. In particular, it has been suggested that in strongly correlated materials, highly excited charge carriers could use their extra energy to excite additional carriers across the Mott gap via impact ionization (II)~\cite{mano.10,co.ma.14}, thus
potentially improving their efficiency beyond the Shockley-Queisser limit~\cite{sh.qu.61}. Although II is also present in conventional semiconductor devices, the time scales for electron-electron scattering are typically much longer than in correlated materials, so that highly excited electrons will mostly dissipate their energy to phonons.

Experimentally, evidence for fast carrier multiplication processes has been detected via pump-probe experiments in VO$_2$~\cite{ho.bi.16}. Oxide heterostructures based on LaVO$_3$/SrTiO$_3$ have been identified as promising candidates, due to the ideal band gap and the strong polar field, able to separate the excited charge carriers~\cite{as.bl.13}. There are however some drawbacks, such as the low mobility of the carriers~\cite{wa.li.15,je.re.18}, which still put into question these materials’ applicability as efficient solar cells. Thus while a large-scale application of Mott-based solar cells may be difficult to achieve, employing them as photodetectors may be more promising on the long run, due to their high photoresponsivity. From the scientific point of view and for future applications it is thus worthwile to further investigate the properties of Mott-based photovoltaic materials and the mechanisms behind their inner workings, to explore whether alternative unexpected paths can be opened. In fact, a lot of theoretical work has been done to understand the II process (see, e.g.~\cite{co.ma.14,ec.we.11,ec.we.13,we.he.14,pe.be.19,so.do.18,ka.wo.20,ma.ev.22}).

As a scattering process, II competes with electron-phonon (e-ph) scattering which is the dominant relaxation mechanism in conventional semiconductors. This may not be the case in Mott insulators, at least in certain cases ~\cite{co.ma.14}. Therefore, it is essential to understand the influence of e-ph scattering on the photocurrent and II in Mott photovoltaic devices, which is the issue we address in this paper.

For this goal we consider the simplified model shown in Fig.~\ref{fig:setup}, consisting of a Hubbard layer located between two leads. The metallic leads have themselves a wide band, which is effectively narrowed due to the weak coupling to intermediate layers which we take as models for the contacts. Electrons on the Hubbard layer interact locally with acoustic phonons. An external periodic electric field induces a Floquet steady state with a current flowing from the lead with the lower to the higher chemical potential, thus transferring its energy.
 We address this periodic problem
 via nonequilibrium Floquet dynamical mean-field theory (F-DMFT).

\begin{figure}[b] 
% \captionsetup{width=\textwidth} 
\includegraphics[width=0.9\linewidth]{./figures_Paper1/setup_voltmeter.pdf}
\caption{(Color online) Schematic representation of the setup. A central, correlated layer with local Hubbard interaction $U$ and e-ph coupling $g$ (green) is sandwiched between two non-interacting layers (red) via the hybridization $v$ (black). The latter are in turn coupled to wide-band fermion reservoirs (blue) with different chemical potentials $\mu_{\text{l}/\text{r}}$, which introduce a broadening $\Gamma=\Gamma_{\text{l}}=\Gamma_{\text{r}}$.}
\label{fig:setup}
\end{figure}

Our goal is to investigate the occurrence of II in this photovoltaic setup and how it is affected by the presence of phonons. 
Specifically we study the behavior of the photovoltaic current, the spectral properties as well as the double occupation as a function of the hybridization strength and the properties of the phonons. Our analysis can only provide qualitative results, while quantitative comparisons with experimental systems require a more realistic setup and are beyond the scope of this work.

We find that a small hybridization to the leads is more favorable for II than a larger one. II therefore plays a crucial role in compensating the smaller charge injection, giving the same order of magnitude in the photocurrent as for the larger hybridization.
In all cases, the interaction with phonons slightly enhances the photocurrent for small driving frequencies, while suppressing it in vicinity of the main peak.

The structure of the paper is as follows: in Sec.~\ref{sec:Model} we describe the model Hamiltonian. Methods, formalism and observables are presented in Sec.~\ref{sec:Method_formalism}. We discuss our results in Sec.~\ref{sec:results} and present our conclusion in Sec.~\ref{sec:conclusions}.

@@

Our setup introduced in Sec.~\ref{sec:intro} and shown in Fig.~\ref{fig:setup} is described by the Hamiltonian
%
\begin{equation}
\label{eq:Hamiltonian}
\begin{split} 
\hat{H}(t)&=\varepsilon_{\text{c}} \sum_{i\sigma}\hat{n}_{i\sigma} -\sum_{\sigma}\sum_{(i,j)} t_{ij}(t) \hat{c}^{\dagger}_{i\sigma} \hat{c}_{j\sigma}
\\ & + U \sum_{i} \hat{n}_{i\uparrow} \hat{n}_{i\downarrow}+\hat{H}_{\text{e-ph}} + \hat{H}_{\text{ph}}+\hat{H}_{\text{lead}}.
\end{split}
\end{equation}
% 
The operator $\hat{c}^{\dagger}_{i\sigma}$ ($\hat{c}_{i\sigma}$) creates (annihilates) a particle with spin $\sigma= \lbrace \uparrow,\downarrow \rbrace$ on the $i$-th lattice site and $\hat{n}_{i\sigma}\equiv \hat{c}^{\dagger}_{i\sigma} \hat{c}_{i\sigma}$ is the particle number operator. We denote with $(i,j)$ the sum over nearest-neighbor sites and with $\varepsilon_{\text{c}} \equiv -U/2$ the on-site energy. The driving field consists of a time-periodic, homogeneous and monochromatic electric field with frequency $\Omega$. In the temporal gauge it enters via the Peierls substitution in the time-dependent hopping in Eq.~\eqref{eq:Hamiltonian}~\cite{peie.33}

\begin{equation}\label{eq:peierls} 
t_{ij}(t) = t_{\text{c}} \ e^{-\ii \frac{q}{\hbar} \left( \vec{r}_j - \vec{r}_i \right) \cdot \vec{A}(t)}, 
\end{equation}
% 
where $t_{\text{c}}$ is the intra-layer hopping, $\vec{A}$(t) the time-dependent vector potential, $\hbar$ Planck's constant and $q$ the charge of the electrons. Following Refs.~\cite{ts.ok.08,mu.we.18}, we choose for simplicity
 $\vec{A}(t)=\vec{e}_{0} A(t)$, which points along the 
lattice body diagonal $\vec{e}_{0}=(1,1,\dots,1)$.  $A(t)=\frac{\hbar}{qa}\aamp\sin(\Omega t)$ with $\aamp=-\frac{qE_0a}{\hbar\Omega}$,  $E_0$ is the  electric field amplitude, and $a$ the lattice spacing ~\cite{ts.ok.08}. 
In this temporal gauge the electric field is $\vec{E}= -\partial_{t}\vec{A}(t) = \vec{e}_{0}E_0 \cos(\Omega t)$.

The local e-ph coupling is included as in Ref.~\cite{ma.ga.22u} with an acoustic phonon branch attached to each lattice site via the Hamiltonian 
\begin{equation}\label{eq:e-ph_ham} 
\hat{H}_{\text{e-ph}} = g \sum_{i\sigma} \hat{n}_{i\sigma} \hat{x}_{i},  
\end{equation}
with $g$ as e-ph coupling and $\hat{x}_{i} \equiv \frac{1}{\sqrt{2}} \left( \hat{b}^{\dagger}_{i} + \hat{b}_i \right)$, where $\hat{b}^{\dagger}_{i}$ ($\hat{b}_{i}$) creates (annihilates) a phonon of the acoustic branch at site $i$, with dispersion relation described by $\hat{H}_{\text{ph}}$~\cite{ma.ga.22u}.  Details on the leads are provided in~\ref{sec:Dyson_equation}.

For mathematical simplicity~\cite{ge.ko.92,ao.ts.14}, 
 we consider the correlated central region as a $d$-dimensional layer and consider the limit of infinite dimensions $d \rightarrow \infty$ by rescaling the hopping as $t_{\text{c}}=t^{\ast}/2\sqrt{d}$. Taking $a=1$, sums over the crystal momentum ${\bf k}$ of a generic quantity $\chi$ read $\sum_{{\bf k}} \chi(\omega,{\bf k}) \rightarrow \int d\epsilon \int   d\overline{\epsilon} \ \rho(\epsilon,\overline{\epsilon}) \chi(\omega;\epsilon,\overline{\epsilon})$ with $\rho(\epsilon,\overline{\epsilon}) = (1/\pi t^{\ast 2}) \exp[-( \epsilon^{2} + \overline{\epsilon}^{2})/t^{\ast 2}]$ as the joint density of states~\cite{ts.ok.08} and with
% 
\begin{align}\label{eq:d-dim_crystal_dep}
\begin{split}  
\epsilon & = -2t_{\text{c}} \sum_{i=1}^{d} \cos(k_i a), \\  
\overline{\epsilon}& = -2t_{\text{c}}\sum_{i=1}^{d} \sin(k_i a). \\
\end{split}  
\end{align}
%

Within the remainder of this paper we set $\hbar = k_{\text{B}} = a = 1 = -q$, $t^{\ast}=2\sqrt{2}$ ~\footnote{With such choice for $t^{\ast}$ we reproduce the DOS of a 2D layer with $t_{\text{c}}=1$  in order to compare with Ref.~\cite{so.do.18}.} and $t_r \equiv \frac{t^{\ast}}{2\sqrt{2}}=1$ as unit of energy.

% In Tommasos paper we also chose -q=1. Given the amplitude of the vector potential above, setting q=-1 allows us to take positive values of E_0 and Omega. same sign of j0 (see old draft). increasing the mu I have negative current meaning that particle/electric current is flowing in the opposite direction.

@@

To correctly describe the nonequilibrium periodic steady state, we use the Floquet generalization of the nonequilibrium Green's function (NEGF) approach~\cite{ts.ok.08}. Every function $G(t,t^{\prime})$ satisfying the periodicity relation $G(t,t^{\prime})=G(t+\tau,t^{\prime}+\tau)$, with $\tau=2\pi/\Omega$ as the period related to the external driving frequency $\Omega$, may be represented as~\cite{ts.ok.08,sc.mo.02u,jo.fr.08}

\begin{equation}
\label{eq:FloquetGF} 
\underline{G}_{mn}(\omega) =\int \dd t_{\text{rel}} \int_{-\tau/2}^{\tau/2} \frac{dt_{\text{av}}}{\tau} \ee^{\ii[\left(\omega+m\Omega\right) t -\left( \omega+n\Omega\right)t^{\prime}]} \underline{G}(t,t'),
\end{equation}

known as the \emph{Keldysh-Floquet} GF. The integration variables $t_{\text{rel}} = t-t^{\prime}$ and $t_{\text{av}} = (t+t^{\prime})/2$ are the relative and average times. Within the rest of this work, we denote a Floquet-represented matrix either as $X_{mn}$ with explicit indices, or use a boldface letter $\bf{X}$. The underline indicates the overall \emph{Keldysh} structure

\begin{equation}\label{eq:Keld-structure}
\underline{{\bf G}} \equiv 
\begin{pmatrix}
{\bf G}^{\text{R}} & {\bf G}^{\text{K}}\\
{\bf 0}         & {\bf G}^{\text{A}} \\
\end{pmatrix},
\end{equation}

which contains the \emph{retarded}, \emph{advanced} and \emph{Keldysh} components $\textbf{G}^{\text{R,A,K}}$, where $\textbf{G}^{\text{A}}=(\textbf{G}^{\text{R}})^{\dagger}$. The \emph{Keldysh} component is defined as $\textbf{G}^{\text{K}} \equiv \textbf{G}^{>} + \textbf{G}^{<}$, with $\textbf{G}^{\lessgtr}$ as \emph{lesser} and \emph{greater} components~\cite{schw.61,keld.65,ra.sm.86,ha.ja}.

@@

\begin{equation}\label{eq:FullDysonEq}
\underline{{\bf G}}^{-1}(\omega_n;\epsilon,\overline{\epsilon}) = \underline{{\bf G}}^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon}) - \underline{{\bf \Sigma}}(\omega_n;\epsilon,\overline{\epsilon}) - \underline{{\bf \Sigma}}_{\text{e-ph}}(\omega_n;\epsilon,\overline{\epsilon}).
\end{equation}

Therein, the lattice GF of the non-interacting part of the Hamiltonian in Eq.~\eqref{eq:Hamiltonian} is

\begin{equation}\label{eq:non-int_InvGF} 
\underline{G}^{-1}_{0,mn}(\omega_n;\epsilon,\overline{\epsilon}) =  \underline{g}^{-1}_{0,mn}(\omega_n;\epsilon,\overline{\epsilon}) - \sum_{\rho\in\{\text{l},\text{r}\}}v^2_{\rho}\underline{g}_{\text{b},\rho}(\omega_n;\epsilon)\delta_{mn}
\end{equation}

with $v_{\text{l}/\text{r}}$ as lead-layer hybridization, $\omega_n \equiv \omega + n\Omega$, $n\in\mathbbm{Z}$, and

\begin{equation}\label{eq:inv_non-int_lat_GF_comps}
\begin{split}
\left[ g^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon})) \right]^{\text{R}}_{mn} & = \left( \omega_n + i0^{+} -\varepsilon_{\text{c}} \right)\delta_{mn} - \varepsilon_{mn}(\epsilon,\overline{\epsilon}), \\
\left[ g^{-1}_{0}(\omega_n;\epsilon,\overline{\epsilon})) \right]^{\text{K}}_{mn} & = 0
\end{split}
\end{equation}

as the noninteracting Green's function of the isolated layer, whose inverse Keldysh component is negligible in the steady state. The Floquet dispersion relation $\varepsilon_{mn}$ for the periodic field in a hyper-cubic lattice is~\cite{ts.ok.08}
